---
title: "Assessing validity of two new taxonomic traits"
subtitle: "Data Analysis"
author: "Allison Fisher"
date: "05/02/2023"
output: html_document
---
After discovering the Papuan genus of microhylid frog, *(Oreophryne)*, is polyphyletic through molecular phylogenetic analysis, I seek to identify morphological characteristics to distinguish the two monophyletic clades. The genus containing type species "*(Oreophryne senckenbergiana)*" will retain the name *(Oreophryne)*. The new genus will be called *(Auparoparo)* in this study.

# Proposed traits

Through visual observation, I have suspected two distinctions between these highly convergent groups. I have measured images of museum specimens (6 species per genus, ~5 individuals per species) using the software ImageJ to assess if these traits are quantitatively different.

One is the ratio of finger pad width to toe pad width. This character will be represented by the variable ft. *(Oreophryne)*s tend to have more equally sized finger and toe pads so I expect their ft ratio to be closer to 1. *(Auparoparo)*s have larger finger pads than toe pads so I expect their ft ratio to be larger than 1.

The other trait I suspect is the shape of the upper palatal grooves. This character will be represented by the variable ap which is the ratio of the width of the grooves clostest to the nares (anterior) vs closest to the pharynx (posterior). *(Oreophryne)*s tend to have more parallel grooves so I expect their ap ratio to be closer to 1. *(Auparoparo)*s have grooves which taper towards the pharynx so I expect their ap ratio to be larger than 1.

# Analysis
I will be testing a variety of analyses to look for significance in these traits.

# Set up for R

```{r }
#| cache = FALSE
knitr::read_chunk('analysiscode.R')
```

Load packages, functions, and paths
```{r, setup}
```
Load data
```{r, loaddata}
```

```{r, functions}
```

# Summarize
Another look at the data

```{r, summarize}
```
```{r, summary.table}
```
Let's look at our variables on a species and genus level
```{r, summarystats}
```


# Box-and-whisker plots
Let's see how the two ratio traits' distributions look across genera

```{r, genus_ft_boxplot}
```

```{r, genus_ap_boxplot}
```
They seem to be different!

# T-tests
Let's see if there's significance via t-test between genera

```{r, ttests}
#| tbl-cap: Oreophryne vs Auparoparo

```

# Phylogeny comparison
I am trimming our phylogeny to only include measured species and then plotting a character heat map using the means for each species
```{r, phylogenyexperiment}
```
