---
title: "Assessing validity of two new taxonomic traits"
subtitle: "Data Processing"
author: "Allison Fisher"
date: "05/02/2023"
output: html_document
---
After discovering the Papuan genus of microhylid frog, *(Oreophryne)*, is polyphyletic through molecular phylogenetic analysis, I seek to identify morphological characteristics to distinguish the two monophyletic clades. The genus containing type species "*(Oreophryne senckenbergiana)*" will retain the name *(Oreophryne)*. The new genus will be called *(Auparoparo)* in this study.

# Proposed traits

Through visual observation, I have suspected two distinctions between these highly convergent groups. I have measured images of museum specimens (6 species per genus, ~5 individuals per species) using the software ImageJ to assess if these traits are quantitatively different.

One is the ratio of finger pad width to toe pad width. This character will be represented by the variable ft. *(Oreophryne)*s tend to have more equally sized finger and toe pads so I expect their ft ratio to be closer to 1. *(Auparoparo)*s have larger finger pads than toe pads so I expect their ft ratio to be larger than 1.

The other trait I suspect is the shape of the upper palatal grooves. This character will be represented by the variable ap which is the ratio of the width of the grooves clostest to the nares (anterior) vs closest to the pharynx (posterior). *(Oreophryne)*s tend to have more parallel grooves so I expect their ap ratio to be closer to 1. *(Auparoparo)*s have grooves which taper towards the pharynx so I expect their ap ratio to be larger than 1.


# Set up for R

```{r }
#| cache = FALSE
knitr::read_chunk('processingcode.R')
```

A variety of packages are needed.

```{r, packages}
#| message: FALSE
#| warning: FALSE
```


```{r, loaddata}
```

# Data Cleanup
Due to import from ImageJ and coversions, I begin with a non-rectangular dataframe. Let's clean it up.


```{r, cleanup}
```

# Checking
Let's check that there's no weird patterns or outliers in our dataset

```{r, exploredata}
```

# Exploratory plots
Let's see how the data looks to begin understanding where we might see patterns

```{r, exploratoryplots}
```
I notice some trends that I would like to test for significance.

# Save data
Now we can save our cleaned up data for analysis

```{r, savedata}
```
